---
title: "Final Key Winter 2024"
---

```{r}
library(tidyverse)
pulsedat <- read_csv("https://tofu.byu.edu/stat230/pulsedat.csv", show_col_types = FALSE) |>
  mutate(across(c("time","diet","id"),factor))
pulsedat
```

```{r}
pulsedat.aov <- aov(pulse ~ diet + Error(id) + time + diet:time, data=pulsedat)
summary(pulsedat.aov)
```

3.  difference between two diets is NOT SIGNIFICANT diet is 0.0869 \> 0.05
4.  NOT significant evidence to reject the default theory diet:time is 0.339 \> 0.05 Insufficient evidence to conclude that there is an interaction.

## Common WRONG answers

```{r}
pulse.CB2 <- lm(pulse ~ id + diet + time + diet:time, data=pulsedat)
anova(pulse.CB2) # note that diet doesn't make it into the ANOVA table since id "soaks up" all the info about diet
```

3.  difference between two diets is inconclusive: diet soaked up by id.
4.  NOT significant evidence to reject the default theory diet:time is 0.3394 \> 0.05 Insufficient evidence to conclude that there is an interaction.

```{r}
pulse.BF2 <- lm(pulse ~ diet + time + diet:time, data=pulsedat)
anova(pulse.BF2)
```

3.  difference between two diets IS significant. diet is 0.01214 \> 0.05
4.  NOT significant evidence to reject the default theory diet:time is 0.60703 \> 0.05 Insufficient evidence to conclude that there is an interaction.
