---
title: "Stat 230 HW#12"
author: "Kimberly Williams"
date: "`r Sys.Date()`"
format: pdf
editor: visual
---

## Numbers 76, 77, 87

```{r, message=FALSE}
library(tidyverse)
```

## 76

#B1 on pages 256-7

a\)

```{r}
dat <- read_table("cow period diet yield
1 1 roughage 608
1 2 partial 716
1 3 full 845
2 1 partial 885
2 2 full 1086
2 3 roughage 711
3 1 full 940
3 2 roughage 766
3 3 partial 832")
cows <- dat %>% mutate(across(c("cow","period","diet"),factor))
cows
```

b\) Average milk yield for each diet

```{r}
tapply(cows$yield,cows$diet,mean)
```

c\) Average milk yield for period

```{r}
tapply(cows$yield,cows$period,mean)
```

Do the data support the idea that milk yields tend to decline over the course of the experiment?

Not really, 796 is about the same as 811

d\) Which differences are bigger: between cows or between periods?

```{r}
tapply(cows$yield,cows$cow,mean)
```

```{r}
print(856-796)
print(894-723)
```

The difference between cows is much different than the difference between periods.

e\) Is the factor of interest for this design observational or experimental? Why is it unusual for an LS design to have the factor of interest be observational?

Experimental. Usually Latin squares are about randomizing the assignments of treatment like the diet.

## 77

```{r}
tapply(cows$yield,cows$diet,mean)
cows.aov <- aov(yield ~ cow + period + diet ,cows)
summary(cows.aov)
```

a\) Does diet have a significant effect on yield? Compare the means for the diets as you interpret the diet effect.

Diet does have a significant effect on yield, roughage averaging the least with a yield of 695, partial averaging the middle with a yield of 811, and full averaging the most with a yield of 957.

b\) Did the nuisance variables (cow and time period) have substantial impact on the yield?

Cow and time period did not have a statistically significant impact on yield, but cow was almost significant with a p-value of 0.0925 only 0.04 bigger than our alpha level of 0.05. Further testing with more subjects is recommended.

## 87

a\) i. Structure of the first of the two squares

|     |     |     |     |
|-----|-----|-----|-----|
| B   | C   | A   | D   |
| C   | D   | B   | A   |
| D   | A   | C   | B   |
| A   | B   | D   | C   |

a\) ii.

yijkl =μ+θi +αj(i) +βk(i) +τl +εijkl, i = 1,2; j = 1,...,4; k = 1,...,4; l = 1,...,4

From left to right, μ is the grand mean; θi is the effect of the ith square (i.e., the effect of the ith rep of the experiment); αj(i) is the effect of the jth operator NESTED in the ith square (i.e., there are a total of 8 different operators); βk(i) is the effect of the kth day NESTED in the ith square (i.e., there are a total of 8 different days); τl is the effect of the lth supplier; and εijkl is the residual.

a\) iii.

| SOURCE           | df  |
|------------------|-----|
| Mean             | 1   |
| Square / "rep"   | 1   |
| Operator / "row" | 6   |
| Day / "column"   | 6   |
| Supplier         | 3   |
| Residual         | 15  |
| TOTAL            | 32  |

a\) iv. Which F distribution would you use to test the effect for Supplier?

F_3,15

a\) v.

```{r}
strengthdat <- read_table("../data/strengthLS.txt", show_col_types = F) |>
  mutate(across(c("rep","operator","day", "supplier"), factor))

strength.aov <- aov(strength ~ rep + operator:rep + day:rep +
         supplier , strengthdat)
# operator nested in square (total of 8 operators)
# day nested in square (total of 8 operators)
summary(strength.aov)
```

b\) For the next part of the problem, assume that the 4 operators in square 1 are the SAME AS the 4 operators in square 2. (That is, operator #2 in square 1 is the same as operator #2 in square 2.) Assume again that the 4 days in square 1 are different from the 4 days in square 2.

b\) i. Same as (a)i.

|     |     |     |     |
|-----|-----|-----|-----|
| B   | C   | A   | D   |
| C   | D   | B   | A   |
| D   | A   | C   | B   |
| A   | B   | D   | C   |

b\) ii.

yijkl =μ+θi +αj +βk(i) +τl +εijkl, i = 1,2; j = 1,...,4; k = 1,...,4; l = 1,...,4

From left to right, μ is the grand mean; θi is the effect of the ith square (i.e., the effect of the ith rep of the experiment); αj is the effect of the jth operator which is CROSSED (not nested) in square (i.e., there are a total of 4 different operators with the same 4 operators showing up as the rows of each of the two squares—there are 4 distinct operator effects instead of the 8 distinct effects in part (a) ); βk(i) is the effect of the kth day NESTED in the ith square (i.e., there are a total of 8 different days); τl is the effect of the lth supplier; and εijkl is the residual.

b\) iii.

| SOURCE           | df  |
|------------------|-----|
| Mean             | 1   |
| Square / "rep"   | 1   |
| Operator / "row" | 3   |
| Day / "column"   | 6   |
| Supplier         | 3   |
| Residual         | 18  |
| TOTAL            | 32  |

b\) iv. Which F distribution would you use to test the effect for Supplier?

F_3,18

b\) v.

```{r}
strength.aov <- aov(strength ~ rep + operator + day:rep + supplier, strengthdat)
# operator crossed with square (total of 4 operators)
# day nested in square (total of 8 operators)
summary(strength.aov)
```
