---
title: "Stat 230 HW#8"
author: "Kimberly Williams"
date: "`r Sys.Date()`"
format: pdf
editor: visual
---

# Numbers 61 and 84

## Problem 61

```{r}
library(tidyverse)
library(DescTools) #PostHocTest, TukeyHSD, ScheffeTest
```

```{r}
#a
ns <- 2:20
powers <- rep(NA,length(ns))
for (i in 1:length(ns)) powers[i] <- power.anova.test(groups=4, between.var=var(c(21,23,25,27)), within.var=4.7^2, sig.level=0.05, n=ns[i])$power

# create plot of n by power
plot(ns, powers,type="l",main="Power Curves",xlab="Sample Size",
               ylab="Power",col="lightblue",lwd=2,ylim=c(0,1))


#c
for (i in 1:length(ns)) powers[i] <- power.anova.test(groups=4, between.var=var(c(27,25,23,21)), within.var=4.7^2, sig.level=0.05, n=ns[i])$power

lines(ns, powers, type="l", col="blue", lty = "dashed", lwd=2,ylim=c(0,1))

#d
for (i in 1:length(ns)) powers[i] <- power.anova.test(groups=4, between.var=var(c(21,21,21,27)), within.var=4.7^2, sig.level=0.05, n=ns[i])$power

lines(ns, powers,type="l",col="darkorange", lty = 4, lwd=2,ylim=c(0,1))

#e
for (i in 1:length(ns)) powers[i] <- power.anova.test(groups=4, between.var=var(c(21,21,27,27)), within.var=4.7^2, sig.level=0.05, n=ns[i])$power

lines(ns, powers,type="l",col="forestgreen", lty = 5, lwd=2,ylim=c(0,1))
```

```{r}
#b
ceiling(power.anova.test(groups=4, between.var=var(c(21,23,25,27)), within.var=4.7^2, sig.level=0.05, power=.85)$n)
```

The smallest group size that gives 85% power is n = 15

## Problem 84

```{r}
headinjury <- read_csv('../data/headinjury.csv')
headinjury$Type <- as.factor(headinjury$Type)

headinjury
```

a)  Name the appropriate design (any of the following are acceptable):

• BF\[1\]\

• One-way ANOVA\

• Completely Randomized Design (CRD)

Model:

yij =μ+αi+εij, i=1,2,...,7 j=1,2,...,10,

where yij is the observed head injury severity scores, μ is the grand mean, αi is the effect for the ith car type, and εij is the error.

b\) Our primary interest is to see if the car types have different mean head-injury severity scores. Write down the appropriate null and alternative hypotheses, carefully defining all symbols.

Ho : α1=α2=α3=α4=α5=α6=α7=0

Ha : At least one αi is different.

c\) Give the ANOVA table and interpret the proper F-test for the hypotheses of interest.

```{r}
#lm <- old way to do it
hi.lm <- lm(HeadInjury ~ Type, data=headinjury)
anova(hi.lm)
```

```{r}
#aov <- use because it's multiple comparisons
hi.aov <- aov(HeadInjury ~ Type, headinjury)
anova(hi.aov)
```

We reject the null hypothesis (p-value = 0.0056). At least one αi is different.

d\) Assume that our primary interest is in constructing a confidence interval for each possible pairwise comparison. If we want to ensure that the family-wise error rate is no greater than 0.05, which multiple-comparison approach is most appropriate? Use your chosen approach and interpret this set of pairwise comparisons—which means are significantly different from each other?

```{r}
PostHocTest(hi.aov, method = "bonferroni")  # wide intervals
```

```{r}
PostHocTest(hi.aov, method = "lsd")  # narrow intervals
```

```{r}
TukeyHSD(hi.aov)  # in between lsd and bonferroni, but still protects from Type 1 errors and "ensures that the family-wise error rate is no greater than 0.05"
```

```{r}
ScheffeTest(hi.aov)  # widest method
```

The proper approach for pairwise comparisons is **Tukey**. Vans are significantly different from medium (p-val = 0.0243419) and heavy (p-val = 0.0059544) cars.

e\) Compare the width of the interval for one of the pairwise comparisons—say, “μvan − μcompact”—when using: (i) Tukey’s HSD, (ii) Scheffe’, and (iii) Bonferroni. Based on the width of the intervals, which is the best approach for these data?

-   Tukey: diff=357.7 (-137.82985, 853.2298) —\> width: 991.0596

-   Scheffe: diff=357.7 (-239.63254, 955.0325) —\> width: 1194.665

-   Bonferroni: diff=357.7 (-157.44049, 872.8405) —\> width: 1030.281

Tukey has the smallest interval and is therefore the best approach
