Calculate the p-vlaue
```{r}
pf(5, 3, 30, lower.tail = FALSE)
```

Calculate power given mus and n
```{r}
mu     <- c(1.2, 2.2, 1.5, 0.9)
sigma2 <- 0.3
power.anova.test(groups = length(mu), n = 8, between.var = var(mu), 
                 within.var = sigma2)
```
Calculate pvalue given F-stat
```{r}
pf(q = 5.23, df1 = 4-1, df2 = 32-4,lower.tail = FALSE)
```

Calculate n given power
```{r}
mu     <- c(1.2, 2.2, 1.5, 0.9)
sigma2 <- 0.3
# find n
power.anova.test(groups = length(mu), between.var = var(mu), 
                 within.var = sigma2, power = 0.95)
```
Import:
```{r}
library(readr)
weightloss_ancova <- read_csv("~/Documents/BYU_Fall_2023/Stat230/data/weightloss (ancova).csv")
weightloss_ancova
```

ANOVA
Visualize
```{r}
stripchart(weightloss ~ treatment, vertical = TRUE, pch = 1, 
           data = weightloss_ancova)

boxplot(weightloss ~ treatment, data = weightloss_ancova)
```

Fit the one-way CRD EFFECTS model with "reference group" constraint
```{r}
fit.plan <- aov(weightloss ~ treatment, data = weightloss_ancova)
alt_model <- lm(weightloss ~ treatment, data = weightloss_ancova)
```
Have a look at the estimated coefficients
```{r}
coef(fit.plan)
coef(alt_model)
```

View all the coefficients
```{r}
dummy.coef(fit.plan)
```

```{r}
library(emmeans)
emmeans(fit.plan, specs = ~ treatment)
```
Change constraints to "sum to zero"
```{r}
options(contrasts = c("contr.sum", "contr.poly")) #for both ordered and unordered values
fit.plan2 <- aov(weightloss ~ treatment, data = weightloss_ancova)
coef(fit.plan2)
dummy.coef(fit.plan2)
```

Generate ANOVA Table
```{r}
summary(fit.plan)
```
Calculate the p-vlaue
```{r}
pf(0.694, 1, 18, lower.tail = FALSE)
```




ANCOVA
```{r}
library(ggplot2) ## This library is needed for the colorful ANCOVA plots
library(rstatix) ## Needed for post-hoc comparisons
library(emmeans) ## Needed for post-hoc comparisons
```
Plot the data
```{r}
boxplot(weightloss ~ treatment, data = weightloss_ancova)
````

```{r}
fit.ancova <- aov(weightloss ~ exercise+treatment, data = weightloss_ancova)
summary(fit.anova)
```

Pairwise comparisons
```{r}
pwc <- emmeans_test(data = weightloss_ancova,
    weightloss ~ treatment, covariate = exercise,
    p.adjust.method = "bonferroni"
  )
pwc
```
```{r}
get_emmeans(pwc)
```

