## Code to Remember
All the libraries I need:
```{r}
library(readr)   ## This will be used to import csv files
library(ggplot2) ## This library is needed for the colorful ANCOVA plots
library(rstatix) ## Needed for post-hoc comparisons
library(emmeans) ## Needed for post-hoc comparisons
```

Calculate the p-vlaue
```{r}
df1 <- 3 #     numerator df or #of groups - 1
df2 <- 30 #    denominator df or (#of groups * n) - # of groups
zscore <- 5 #  z-score or z-ratio. MSTreatment/MSError = MST/MSE

pf(q = zscore, df1 = df1, df2 = df2, lower.tail = FALSE)
pf(5, 3, 30, lower.tail = FALSE)
```

Calculate power given mus and n
```{r}
mu     <- c(1.2, 2.2, 1.5, 0.9) # list of mu_i
sigma2 <- 0.3   # this is literally just sigma^2 or s^2 if you only have a sample pop.
power.anova.test(groups = length(mu), n = 8, between.var = var(mu), 
                 within.var = sigma2)
```

Calculate pvalue given F-stat - example 2
```{r}
pf(q = 5.23, df1 = 4-1, df2 = 32-4,lower.tail = FALSE)
```

Calculate n given power
```{r}
mu     <- c(1.2, 2.2, 1.5, 0.9)
sigma2 <- 0.3
# find n
power.anova.test(groups = length(mu), between.var = var(mu), 
                 within.var = sigma2, power = 0.95)
```


Import:
```{r}
data_set <- read_csv("~/Documents/BYU_Fall_2023/Stat230/data/"___".csv")
data_set
```
define varibles:
```{r}
continuous <- y_or_response
categorical <- x_or_explanetory
xi <- covariant
```

ANOVA
Visualize
```{r}
stripchart(continuous ~ categorical, vertical = TRUE, pch = 1, 
           data = data_set)

boxplot(continuous ~ categorical, data = data_set)
```

Fit the one-way CRD EFFECTS model with "reference group" constraint
```{r}
fit.plan <- aov(continuous ~ categorical, data = data_set)
alt_model <- lm(continuous ~ categorical, data = data_set)
```
Have a look at the estimated coefficients
```{r}
coef(fit.plan)
coef(alt_model)
```

View all the coefficients
```{r}
dummy.coef(fit.plan)
```

```{r}
emmeans(fit.plan, specs = ~ categorical)
```
Change constraints to "sum to zero"
```{r}
options(contrasts = c("contr.sum", "contr.poly")) #for both ordered and unordered values
fit.plan2 <- aov(continuous ~ categorical, data = data_set)
coef(fit.plan2)
dummy.coef(fit.plan2)
```

Generate ANOVA Table
```{r}
summary(fit.plan)
```

ANCOVA
Plot the data
```{r}
boxplot(continuous ~ categorical, data = data_set)
```

```{r}
fit.ancova <- aov(continuous ~ xi+categorical, data = data_set)
summary(fit.ancova)
```

Pairwise comparisons
```{r}
pwc <- emmeans_test(data = data_set,
    continuous ~ categorical, covariate = xi,
    p.adjust.method = "bonferroni"
  )
pwc
```
```{r}
get_emmeans(pwc)
```



t.test
```{r}
t.test(x, y = NULL,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95, ...)
```

Notes:
SE --> Standard Error --> s/sqrt(n)
p-value -->
t-test --> (x_bar - mu)/sqrt(sigma^2 / sqrt(n))
conf int --> x_bar +- t*SE
test_stat --> x_bar / SE
power --> probability you reject the null when the null is wrong
      --> 1 - probability of a Type 2 Error

```{r}     
sigma <- 5/sqrt(15)
z_score <- qnorm(.05)
sample_mean <- 30
critical_z <- sample_mean + z_score*sigma
pnorm(critical_z, 28.5, sigma) # --> probability under a normal dist.
# pnorm(x, mu, sigma) --> probability of this point or less
```

test stat for test comparing two contents:
```{r}
x_bar_1 <- "sample mean 1"
x_bar_2 <- "sample mean 2"
n_1 <- "number in sample 1"
n_2 <- "number in sample 2"
sd_1 <- "standard deviation 1"
sd_2 <- "standard deviation 2"
s_pool <- ((n_1-1)*sd_1+(n_2-1)*sd_2)/(n_1+n_2-2)
s_sqr_pool <- s_pool*s_pool
t_stat <- (x_bar_1 - x_bar_2) / sqrt(s_sqr_pool*(1/n_1 + 1/n_2))
```

paired t.test
```{r}
x <- c(1.3,2.5,3.1,4.3,5.4)
y <- c(2.3,3.8,4.6,5.5,6.4)
t.test(x,y,paired=TRUE)
```



